---
title: "Data Overview - Univariate"
output:
  html_document:
    df_print: paged
---

## 1. Import relevant libraries

```{r}
library(ArrowDQAToolkit)
source("helpers.R")
```


--- 
## 2. Loading data set
Set up path to load data

```{r load data}
data_path <- file.path("/.", "cluster_data", "vrdata", "raw", "parquet")
raw_data <- open_dataset(data_path)
```

Glimpse data 

```{r Glimpse data}
print(n=26,
      raw_data %>% 
        glimpse())
```


--- 
## 3. Overview of data 

### 3.1. Numeric variables

**tetanus mom** - Number of tetanus shots the child's mom has taken
```{r tetanus_mom}
# --- Number of entries grouped by tetanus_mom --- 
 raw_data %>% 
        group_by(tetanus_mom) %>% 
        count() %>% 
        arrange(desc(tetanus_mom)) %>% 
        collect()

# --- Basic stats --- 
raw_data %>% 
        arrow_summary("tetanus_mom") %>% 
        collect()
```
Observations:
- Positively skewed, with extreme values (outliers)

**vacorder** - Vaccination order

```{r vacorder}
# --- Number of entries grouped by vacorder --- 
raw_data %>% 
        group_by(vacorder) %>% 
        count() %>% 
        arrange(desc(vacorder)) %>% 
        collect()

# --- Basic stats --- 
raw_data %>% 
        arrow_summary("vacorder") %>% 
        collect()
```
Observations:
- Positively skewed, with extreme values (outliers)

**age**
compute new column to analyze age of children
```{r age}

# create new R dataframe containing vaccine name and age (of vaccination)
raw_data %>%
  dplyr::filter(dob <= vacdate) %>%
  mutate(
    vacname = vacname,
    week_age = arrow_weeks_between(dob, vacdate),
    .keep="none") %>%
  arrow_summary("week_age") %>% 
  collect()
```

### 3.2. Categorical variables
**sex** 
```{r sex}
raw_data %>%
  group_by(sex) %>% 
  count() %>% 
  collect()
```


**vacname** - Vaccine name
```{r vacname}
# number of NA entries
raw_data %>% 
  dplyr::filter(is.na(vacname)) %>% 
  count() %>% 
  collect()

# vaccine name
raw_data %>% 
        group_by(vacname) %>% 
        count(sort = TRUE) %>% 
        collect()

```
Observations:
- Non-standardized naming schemes e.g. Uon van and TETANUS vaccine, Teatanus antitoxin vs  TETANUS antitoxin
- Typos

**vactype** - Type of vaccination
Label:
TCMR (Tiem chung mo rong) - Expanded program on Immunization
TCDV (Tiem chung dich vu) - Vaccination through services 
TCCD (Tiem chung chien dich) - Vaccination campaign
```{r vactype}
raw_data %>% 
        group_by(vactype) %>% 
        count() %>% 
        collect()
```

**vacplace_type** - Type of vaccination place
TCMR (Tiem chung mo rong) - EPI
DV (Dich vu) -  service 
BV (Benh vien) - hospital
Dia diem khac - others
```{r vacplace_type}
raw_data %>% 
        count(vacplace_type) %>% 
        collect()
```

**vacplace**
```{r vacplace}
raw_data %>% 
        group_by(vacplace) %>% 
        count(sort=TRUE) %>% 
        collect()
```

**vacplace0**
```{r vacplace0}
raw_data %>% 
        group_by(vacplace0) %>% 
        count(sort=TRUE) %>% 
        collect()
```
Observations:
- Inconsistent categories (type of vacplace vs name of vacplace)

**place_update**
```{r place_update}
raw_data %>% 
        group_by(place_update) %>% 
        count(sort=TRUE) %>% 
        collect()

# check whether place update the same as vacplace (if returns 0 row -> vacplace the same as place update)
raw_data %>% 
  dplyr::filter(place_update != vacplace) %>% 
  collect()
```
Observation:
- Vacplace values are exactly the same as that of place_update


**ethnic**
```{r ethnic}
 raw_data %>% 
        group_by(ethnic) %>% 
        count(sort=TRUE) %>% 
        collect()
```
**province**
```{r province}
raw_data %>%
        group_by(province) %>%
        count(sort=TRUE) %>% 
        collect()
```
Observation:
- Expected 63, got 64 provinces (Tinh tap huan)

**province_reg** - Province registered for vaccination (?)
```{r province_reg}
raw_data %>%
        group_by(province_reg) %>%
        count(sort=TRUE) %>% 
        collect()
```
Observation:
- Expected 63, got 62 (no entries for Binh Dinh)

### 3.3. Dates 

**dob** - Date of birth 
```{r dob}
# Number of NA entries
raw_data %>% 
  dplyr::filter(is.na(dob)) %>% count() %>% collect()

# Basic stats
raw_data %>% 
  summarize(
            min_value = min(dob, na.rm=TRUE),
            max_value = max(dob, na.rm=TRUE),
          ) %>% 
  collect()
```

**vacdate** - Date of vaccination
```{r vacdate}
# number of NA entries
raw_data %>% 
  dplyr::filter(is.na(vacdate)) %>% 
  count() %>% 
  collect()

# summary of vacdate
raw_data %>% 
  summarize(
    min_date = min(vacdate, na.rm = TRUE),
    max_date = max(vacdate, na.rm = TRUE)
  ) %>% collect()

# entries where date of vaccination is before date of birth
raw_data %>% 
  dplyr::filter(vacdate < dob) %>% 
  select(vacdate, dob, place_update, file, rn) %>% 
  collect()

# entries where date of vaccination is before date of birth, grouped by place update
raw_data %>% 
  dplyr::filter(vacdate < dob) %>% 
  count(place_update, sort=TRUE) %>% 
  collect()

```
Observations:
- Extreme values way over the plausible range (1-01-01 to 9202-08-21)
- 115484 entries where dob is less than vacdate (contradictory values)

### 3.4. String variables 
```{r pid}
# number of NA entries
raw_data %>% 
  dplyr::filter(is.na(pid)) %>% 
  count() %>% 
  collect()

# Number of records for each pid/child
raw_data %>% 
  count(pid, sort=TRUE) %>% 
  collect()

# Summarize number of records for each pid
raw_data %>% 
  count(pid) %>% 
    arrow_summary("n") %>% 
    collect()
```
Observation:
- Approximately 14,405,018 unique pid entries
- Some pids has number of records above plausible range

**caregiver** 
```{r caregiver}
# Check if rows have punctuation characters
raw_data %>% 
  dplyr::filter(str_detect(caregiver, "[[:punct:]]+")) %>% 
  select(pid, caregiver) %>% 
  distinct() %>% 
  collect()
```

**name**
```{r name}
# Check if rows have punctuation characters
raw_data %>% 
  dplyr::filter(str_detect(name, "[[:punct:]]+")) %>%
  select(pid, name) %>% 
  distinct() %>% 
  collect()

# --- Extract rows with disabled (?) pid 
raw_data %>% 
  dplyr::filter(str_detect(name, stringr::regex("(Huỷ|Trùng)", ignore_case=TRUE))) %>% 
  collect()

# --- Extract rows with practically no name
raw_data %>% 
  dplyr::filter(name == ".") %>% 
  count(pid) %>% 
  collect()
```




